---
title: "Machine Learning Project"
output: html_document
---

## Summary

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.

The goal of this project is to predict the style subjects did the exercise, stated in the “classe” variable, using the other variables.

## Downloading data

```{r}

url_training <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
url_testing <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

download.file(url_training, destfile = "training.csv", method = "curl")
download.file(url_testing, destfile = "testing.csv", method = "curl")

training <- read.csv("training.csv", row.names = 1, na.strings=c("NA","#DIV/0!",""))
testing <- read.csv("testing.csv", row.names = 1, na.strings=c("NA","#DIV/0!",""))

```


## Cleaning data

```{r}

library(caret)
# Drop the first 5 columns
training_clean <- training[,6:length(colnames(training))]
testing_clean <- testing[,6:length(colnames(testing))]

# Drop colums with NA
training_clean <- training_clean[, colSums(is.na(training_clean)) == 0] 
testing_clean <- testing_clean[, colSums(is.na(testing_clean)) == 0]

# Exclude near zero variance predictors
nzv <- nearZeroVar(training_clean,saveMetrics=TRUE)
training_clean <- training_clean[, nzv$nzv==FALSE]

training_clean$classe = factor(training_clean$classe)
```

##Splitting data

```{r}

set.seed(1992)
in_training <- createDataPartition(training_clean$classe, p=0.75, list=FALSE)
training_split <- training_clean[in_training, ]
validating_split <- training_clean[-in_training, ]

```


##Training the model

```{r}

library(randomForest)
modelrf <- train(classe ~ ., data=training_split, method="rf")

```


##Validating the model

```{r}

pred <- predict(modelrf, validating_split)
confusionMatrix(pred, validating_split$classe)

```

The accuracy of this model is 0.9988.


## Testing the model

```{r}

ptest <- predict(modelrf, testing)
ptest

```